---
title: "Final"
author: "Ade Olu-Ajeigbe"
date: "5/3/2021"
output: pdf_document
---

1. Redefine the “quality” variable as a factor with 2 levels-‘high’ (Quality score 6 and
above) and ‘low’ (Quality score 5 or less) Hint: One way to do this is using the ifelse
function in R. [Make sure you keep this binary variable only for quality and not the
continuous one]
(a) (4 points) To study variables that affect the quality of wine, fit a logistic regression
model with all covariates.
(b) (5 points) Interpret the coefficients for significant covariates (How does the odds of
high quality wine change with these covariates?).
(c) (4 points) Reduce the model (reduce the number of covariates) to fit the best possible
model using all 1500 observations.
(d) (4 points) Is there statistical evidence from your reduced model that % of alcohol is
associated with an increase in the probability of high quality wine? Explain.
(e) (2 points) Explain why linear regression is inappropriate for modeling the prevalence
of high quality wine. When could you use linear regression instead?

```{r, echo = TRUE}


data <- read.csv(file = "https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequaliy-white.csv", header = TRUE, sep = ";")
final<-data2[sample(nrow(data2), 1500), ]


final$quality <- ifelse(final$quality > 6, "high", as.character(final$quality))

final$quality <- ifelse(final$quality < 5, "low", as.character(final$quality))

final$quality <- final$quality == "high" 

head(final)



#PartA)
cols<-c(1:11) #columns numbers 1 through 11
final[cols]<-sapply(final[cols], as.numeric)#changes first 11 columns to numeric


model<-glm(quality ~ .,family = binomial(logit), data=final)
summary(model)

# residual sugar, volatile.acidity,density, pH, sulphates, and fixed.acidity are the significant values from least to greatest

#Part B)
#residual sugars were the most significant since they had the lowest p-values

exp(model$coefficients)

#Part C)
model2<-glm(quality ~ residual.sugar + volatile.acidity + pH + density + sulphates + fixed.acidity, family = binomial(logit), data =final)
summary(model2)

#Part D)
#The alcohol seems to not have an intense correlation between the two values present

#Part E)
#Linear regerrions would need values that are more closely related that can be interpreted fully.
```

(a) (10 points) Perform k-fold cross validation with your model from 1(c) with k=10.
Report the accuracy of your model.
(b) (10 points) Report the accuracy of your model through leave one out cross validation.
For this dataset, which method (k-fold or leave one out) would you recommend?


```{r, echo = TRUE}
x <- final[sample(1:nrow(final)),]
library(caret)
library(e1071)

#set.seed(134)
Train <- createDataPartition(final$quality, p=0.75, list=FALSE)
training <- x[ Train, ]
testing <- x[ -Train, ]


train.control <- trainControl(method = "repeatedcv", number = 10, savePredictions = TRUE, repeats = 5)


training$quality<-factor(training$quality)



kfoldtestss<- train(quality ~ residual.sugar + volatile.acidity + pH + density + sulphates + fixed.acidity, data = training, trControl = train.control, method = "glm", family = "binomial")
print(kfoldtestss)

#Part B

acc <- NULL
for(i in 1:nrow(final))
{
train <- final[-i,]
test <- final[i,]
# Fitting
loo_m <- glm(quality ~ residual.sugar + volatile.acidity + pH + density + sulphates + fixed.acidity, family =binomial(logit), data = train)
# Predict results
pred <- predict(loo_m,newdata=test,type="response")
# If prob > 0.5 then 1, else 0
results <- ifelse(pred > 0.5,1,0)
# Actual answers
answers <- test$quality
# Calculate accuracy
misClasificError <- mean(answers != results)
# Collecting results
acc[i] <- 1-misClasificError
}
mean(acc)



```


3. Swarnali can have 0, 1 or 2 ice-creams in a day depending on whether it is a Hot or Cold
day. Consider the following Hidden Markov Model: 
P(0|H) = 0.2
P(1|H) = 0.5
P(2|H) = 0.3
P(0|C) = 0.6
P(1|C) = 0.3
P(2|C) = 0.1

hot to cold = 0.3
cold to hot = 0.4
hot to hot = 0.7
cold to cold = 0.6

(a) (10 points) On three consecutive days, number of ice-creams consumed were 1 0 2. If
all three days had similar weather, which of HHH and CCC is more likely?
(b) (10 points) Use Viterbi algorithm to find the maximum probability path and the
maximum probability of this sequence.

```{r, echo = TRUE}
#Part A
HHH <- (0.5 *0.5)*(0.7* 0.2)*(0.7 * 0.3)
HHH
CCC <- (0.5 *0.3)*(0.6* 0.6)*(0.6 *0.1)
CCC
##hot is more likely

#Part B
markov <- function(x,P,n){ seq <- x
for(k in 1:(n-1)){
seq[k+1] <- sample(x, 1, replace=TRUE, P[seq[k],])}
return(seq)
}
hmmdat <- function(A,E,n){
observationset <- c(1:3) #3 diferent days
hiddenset <- c(1,2) #hot denoted by 1, cold denoted by 2
x <- h <- matrix(NA,nr=n,nc=1)
h[1]<-1
x[1]<-sample(observationset,1,replace=TRUE,E[h[1],])
h <- markov(hiddenset,A,n)
for(k in 1:(n-1)){x[k+1] <- sample(observationset,1,replace=TRUE,E[h[k],])}
out <- matrix(c(x,h),nrow=n,ncol=2,byrow=FALSE)
return(out)
}

E <- matrix(c(0.2,0.5,0.3,0.6,0.3,0.1),2,3,byrow=TRUE) #emission matrix
E

A <- matrix(c(0.7,0.3,0.6,0.4),2,2,byrow=TRUE) #transition matrix
A

dat <- hmmdat(A,E,100)
colnames(dat) <- c("observation","hidden_state")
rownames(dat) <- 1:100
t(dat)

viterbi <- function(A,E,x) {
v <- matrix(NA, nr=length(x), nc=dim(A)[1])
v[1,] <- 0; v[1,1] <- 1
for(i in 2:length(x)) {
for (l in 1:dim(A)[1]) {v[i,l] <- E[l,x[i]] * max(v[(i-1),] * A[l,])}
}
return(v)
}
vit <- viterbi(A,E,dat[,1]) #using algorithm
vitrowmax <- apply(vit, 1, function(x) which.max(x)) #tracing back max prob p
datt <- cbind(dat,vitrowmax)
colnames(datt) <- c("observation","hidden_state","predicted state")
t(datt)

hiddenstate <- dat[,2]
tab<-table(hiddenstate, vitrowmax)
accuracy<-sum(diag(tab))/sum(tab)
misClasificerror<-1-accuracy
misClasificerror




```
4.This question has two related parts.
(a) (10 points) Consider the 4 nucleotides of a DNA sequence ‘a’, ‘c’, ‘g’, ‘t’. For these 4
states, keeping in mind properties of a transition matrix, define any 4 by 4 transition
matrix of your choice. Now, using this transition matrix of your choice, P, and an
initial probability vector 𝜋 = (0.4,0.1,0.1,0.4), generate a Markov chain of length
1000.
(b) (10 points) For the chain generated in (a), use a chi-square goodness of fit test to
check if purines (‘a’, ‘g’) and pyrimidines (‘c’, ‘t’) have equal probability. 
```{r, echo=TRUE}

library(ade4)
library(seqinr)
x<-read.fasta("C:/Users/willi/OneDrive/Documents/E.Coli/AE005174v2.fas")
ecoli<-c(x[[1]],x[[2]])
P<-prop.table(table(ecoli[4000:4999],ecoli[4001:5000]),1)
P #Transition matrix

pi<-c(0.4,0.1,0.1,0.4)
nucleotides <- c("a","c","t","g")
length<-1000
chain<-rep(NA,length)
chain[1]<-sample(nucleotides,1,p=pi)
chain[1]
for (i in 1:999){
 chain[i+1]<-sample(nucleotides,1,p=P[chain[i],])
 }
head(chain)

#b)
x<-table(chain[1:1000])[c("a","c","t","g")]
x<-c(2*(x[4]+x[2]), 2*(x[1]+x[3]))#freq of G, #freq of others
p<-c(0.5,0.5)
chisq.test(x,p=p)




```



6. (a) (5 points) For your data set ‘final’, perform principal component analysis on the 11
predictor variables only. (leave out quality of wine for this part)
(b) (2+2+2 points) To retain at least 90% variation in the data, how many principal
components (PCs) should be retained? How much variation are you retaining exactly
here? Justify why you see so many or so few PCs to retain 90% variation.
(c) (3 points) Express the first PC as a linear combination of the 11 original variables.
Comment on these coefficients.
(d) (2 points) Should there be any correlation between PC1 and PC2? Justify.
(e) (4 points) Plot PC1 vs PC2 and overlap quality of wine (‘high’ and ‘low’) with
different colors. Comment.


```{r, echo = TRUE}

colnames(quality)<-NULL
pca<- prcomp(final, center = TRUE,scale. = TRUE)
#standardization is required for PCA
#So let us use the correlation matrix instead of the variance-covariance matrix
R<-cor(t(final))
dim(R)
E<-eigen(R) 

values<-E$values
sum(values[1:52])/sum(values)
pca <- prcomp(data, center = TRUE,scale. = TRUE)

print(pca) #s
summary(pca) 
plot(pca,type="l")
#BIPLOT
library(ggfortify)
autoplot(pca, data = iris, loadings = TRUE, loadings.label=TRUE)
#is there clustering with respect to Species?
autoplot(pca, data = final, loadings = TRUE, loadings.label=TRUE, colour='quality')

biplot(pca)



```


